---
title: "STAT 650 Homework 2"
author: 'YiChia Wu(UIN: 132006360)'
output:
  html_document:
    df_print: paged
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("foreign")
ambulance <- read.csv("./ambulance.csv")
student <- read.dta(file = "./student.dta")
library(dplyr) 
library(knitr)
library("gridExtra")
library(ggplot2)
library(e1071) 
```

## Problem 1
a. What fraction of the calls are cancelled? \
```{r}
a <- subset(ambulance, Precancel_Dur != '-1')

length(a$Precancel_Dur) / length(ambulance$Precancel_Dur)
```
b. Provide a histogram for the time in seconds until cancellation for all of those calls that are eventually
cancelled. \
Convert the vector x of values to second.
```{r}
aaselect <- select(a, Precancel_Dur)
aaselect$Precancel_Dur <- aaselect$Precancel_Dur *24 *60 *60

knitr::kable(head(aaselect))
```

  + making the full plot: 
```{r}
hist(x=aaselect$Precancel_Dur, main="Histogram of Precancel Dur", xlab="Time(second)", ylab="Frequency") 
```

  + making the plot of values $\leq$ 900 seconds: 
```{r}
bfilter<- filter(aaselect, Precancel_Dur<=900)

hist(x=bfilter$Precancel_Dur, main="Histogram of Precancel Dur (less than 900s)", xlab="Time(second)", ylab="Frequency") 
```

  + How many (percentage) durations are > 900 seconds? 
```{r}
(length(aaselect$Precancel_Dur) - length(bfilter$Precancel_Dur)) / length(aaselect$Precancel_Dur) * 100 
```

c. What fraction of these calls require transport to a
hospital? \ 
```{r}
c <- subset(ambulance, Precancel_Dur == '-1')
cc <- subset(c, Time_Arrive_Hosp != '-1')
cc <- select(cc, Precancel_Dur, Time_Arrive_Hosp)
knitr::kable(head(cc))
```
  + answer: 
```{r}
length(cc$Precancel_Dur) / length(c$Precancel_Dur) 
```

d. What are the mean and median times spent at the scene for calls that require transport the hospital? \
  + Mean: 1209.944 (second)
  + Median: 1076.840 (second)
```{r}
d <- subset(c, Time_Arrive_Hosp != '-1')
d <- select(d, Time_Arrive_Hosp, Time_Vehicle_At_Scene, Time_Depart_Scene)

d$Time_Vehicle_At_Scene <- d$Time_Vehicle_At_Scene *24 *60 *60
d$Time_Depart_Scene <- d$Time_Depart_Scene *24 *60 *60

d <- mutate(d,spent=Time_Depart_Scene-Time_Vehicle_At_Scene)
d <- select(d, spent)

summary(d)
```
e. Generate a scatter plot of the time spent at scene versus the time spent at hospital (calls
required transport to the hospital) \
```{r}
d <- subset(c, Time_Arrive_Hosp != '-1')
d <- select(d, Time_Vehicle_At_Scene, Time_Depart_Scene, Time_at_Hospital)

d$Time_Vehicle_At_Scene <- d$Time_Vehicle_At_Scene *24 *60 *60
d$Time_Depart_Scene <- d$Time_Depart_Scene *24 *60 *60
d$Time_at_Hospital <- d$Time_at_Hospital *24 *60 *60

d <- mutate(d,spent=Time_Depart_Scene-Time_Vehicle_At_Scene)
d <- select(d, spent, Time_at_Hospital)

plot(x=d$spent,            
     y=d$Time_at_Hospital,
     main="Time spent at scene v.s Time spent at hospital",
     xlab="At scene(second)",           
     ylab="At hospital(second)")   
```
\
From the plot I found: \

+ Most of the data spent time at scene and hospital less than 4000 seconds(about one hour.) In specific, it's highly to spend time at scene less than 2000 seconds.
+ It's unusual to spend lots of time at scene more than 8000 seconds(about two hour.)
+ When spending more than 6000 seconds at scene, they usually spent less than 4000 seconds at hospital.
+ Without outliers, it's usual to spend less than 8000 seconds at hospital.

f. What fraction of non-cancelled calls
in this dataset have response times under 10 minutes? \
10 minutes = 600 seconds \
To find fraction that less than 600 seconds: 
```{r}
f <- subset(c, Time_Arrive_Hosp != '-1')
f <- select(f, Time_Rec, Time_Vehicle_At_Scene)

f$Time_Vehicle_At_Scene <- f$Time_Vehicle_At_Scene *24 *60 *60
f$Time_Rec <- f$Time_Rec *24 *60 *60

f <- mutate(f,spent=Time_Vehicle_At_Scene-Time_Rec)
f <- select(f, spent)

ffilter<- filter(f, spent<600)
length(ffilter$spent) / length(f$spent)
```
\

## Problem 2 (The game of craps)
a. Point $= 4$ \
For each roll, P(win the game)$= \frac{3}{36}$, P(lose the game)$= \frac{6}{36}$ \
\
By observing the probability in each roll: \
P(win the game in roll 2) $= \frac{3}{36}$ \
P(win the game in roll 3) $= \frac{3}{36} \cdot \frac{27}{36}$ \
P(win the game in roll 4) $= \frac{3}{36} \cdot \frac{27}{36}\cdot \frac{27}{36}$ \
\quad It will be a infinite Sum of geometric sequence to calculate the answer. \
\
By sum to infinity formula, with $a_{1} = \frac{3}{36}$, $r = \frac{27}{36}$, \
We can get: \
P(win) =   $\frac{a_{1}}{1-r}$= $\frac{3/36}{9/36}$  = $\frac{1}{3}$ \
\
In other words, the answer can also be calculated by $\frac{P(roll\ a\ 4)}{P(roll\ a\ 4) + P(roll\ a\ 7)}$

b. Calculate $p_{i}$, $i = 2\ to\ 12$ \
Same as (a). Calculate each in same way.\
\
First, address special case: 
  + $p_{2}$, $p_{3}$, $p_{12}$ will lose in first roll (with $\frac{4}{36}$ to get 0)
  + $p_{7}$, $p_{11}$ will win in first roll.(with $\frac{8}{36}$ to get 1)\
\
Other case: 
  + $p_{4}$ = $p_{10}$ = $\frac{1}{3}$
  + $p_{5}$ = $p_{9}$ = $\frac{4/36}{10/36}$ = $\frac{2}{5}$
  + $p_{6}$ = $p_{8}$ = $\frac{5}{11}$\

c. Compute P(Win)\
By summarize all above in (b), we can get the answer.\
P(Win) = $1\cdot\frac{8}{36} + \frac{1}{3}\cdot\frac{3}{36}\cdot2 + \frac{2}{5}\cdot\frac{4}{36}\cdot2 + \frac{5}{11}\cdot\frac{5}{36}\cdot2 + 0\cdot\frac{4}{36}$ \
= $\frac{244}{495}$


## Problem 3 naive Bayes classifier
$P(play\ golf\ | sunny, hot, high, windy)$ \
=$\frac{P(sunny| play\ golf)P(hot| play\ golf)P(high| play\ golf)P(windy| play\ golf)P(play\ golf)}{P(sunny, hot, high, windy)}$\
\
Calculate each feature: \

Outlook_Play Golf | Yes | No
:------------- | :-------------: | :-------------:
sunny  | 3/9 | 2/5
overcast  | 4/9 | 0
rainy | 2/9 | 3/5

Temp_Play Golf | Yes | No
:------------- | :-------------: | :-------------:
Hot | 2/9 | 2/5
Mild  | 4/9 | 2/5
Cool | 3/9 | 1/5

humidity_Play Golf | Yes | No
:------------- | :-------------: | :-------------:
high  | 3/9 | 4/5
normal  | 6/9 | 1/5

Windy_Play Golf | Yes | No
:------------- | :-------------: | :-------------:
True  | 3/9 | 3/5
False  | 6/9 | 2/5

Play Golf | Yes | No
:------------- | :-------------: | :-------------:
Play Golf  | 9/14 | 5/14

Therefore, we can get answer: \
$P(play\ golf\ | sunny, hot, high, windy)$ \
= $\frac{3/9 \cdot 2/9 \cdot 3/9 \cdot 3/9 \cdot 9/14}{3/9 \cdot 2/9 \cdot 3/9 \cdot 3/9 \cdot 9/14 + 2/5 \cdot 2/5 \cdot 4/5 \cdot 3/5 \cdot 5/14}$ \
$\approx$ $\frac{0.0053}{0.0053 + 0.0274}$ $\approx$ $0.1621$ 

## Problem 4 naive Bayes classifier
a. Plot the side by side boxplots for **math** and **science** scores with respect to different program choices using the package
ggplot2. \
```{r}
a1 <- ggplot ( data = student , aes(x = prog , y = math , color = prog )) + geom_boxplot()+ ylab ("math")+ xlab ("prog") + ggtitle("math score boxplot")
a2 <- ggplot ( data = student , aes(x = prog , y = science , color = prog )) + geom_boxplot()+ ylab ("science")+ xlab ("prog") + ggtitle("science score boxplot")

grid.arrange(a1, a2, nrow = 1, ncol = 2)
```


b. Fit the Naive Bayes to classify the program choice prog using the three categorical variables gender **female**, school type **schtyp** and social economic status **ses**. \
```{r}
student_sub <- select(student,female, schtyp, ses, prog)

classifier <- naiveBayes(prog ~ ., data = student_sub )
classifier
```
(a) What is the fraction of the students who choose the academic program? \
academic program: 0.525

(b) For a female student, what is the probability that she chooses the vocation program? \
P(vocation | female) = 0.2477064
```{r}
student_sub <- select(student,female, prog)
classifier <- naiveBayes(female ~ ., data = student_sub )
classifier
```

(c) Are the "social economic status" and "whether the student chooses the academic program" independent from each other? \
```{r}
student_sub <- select(student,ses, prog)
classifier <- naiveBayes(ses ~ ., data = student_sub )
classifier
```
**No.** If "social economic status" and "whether the student chooses the academic program" are independent from each other, \
P(ses and academic) = $P(ses) \cdot P(academic)$  or P(ses |  academic) = P(ses)\
\
However, by checking the result with P(ses=low | academic): \
P(ses=low | academic) = 0.1809524 \
P(ses=low) = 0.235 \
It's not match for these two probability. And so do other features.\
Therefore, "social economic status" and "whether the student chooses the academic program" are **not independent** from each other.


