---
title: "STAT 650 Homework 3"
author: 'YiChia Wu(UIN: 132006360)'
output:
  html_document:
    df_print: paged
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("foreign")
library(dplyr) 
library(knitr)
library("BiocManager")
BiocManager::install("Rgraphviz")
library(Rgraphviz)
library(bnlearn)
```

## Problem 1
a. **What is the change that a bag identified as containing explosives actually does contain explosives?** \

+ P(identified as explosive) = P(actually contain & identified as explosive) + P(without contain & identified as explosive)\
= $\frac{10}{3000000}\cdot0.95 + (1-\frac{10}{3000000}){3000000}\cdot0.005$ \
= 0.0000031666... + 0.004999... = 0.00500315\
**Answer:**\
$\frac{0.0000031666}{0.00500315} = 0.000633$

b. **If we wanted this probability to be at least 0.5. then what should the probability of correctly identifying a bag not containing explosives need to be?** \
Assume: P(without contain & identified as not explosive) = p \
Equation will be: \
$\frac{0.0000031666}{0.0000031666 + (1-10/300000)(1-p)} = 0.5$ \
1-p = 0.000003166 \
**Answer:**\
p = 0.9999968

c. **Is it possible to obtain a probability of 0.5 of correctly identifying a bag as containing explosives by instead increasing the chance of correctly identifying bags containing explosives? Explain your answer.** \
**Answer: No.**\
Assume: new probability of correctly identifying bags containing explosives = p \
$P(identified\ as\ explosive)  = 0.5$ \
$\frac{10}{3000000}\cdot p + (1-\frac{10}{3000000}){3000000}\cdot0.005 = 0.5$ \
Only when p = 135000, it may obtain a probability of 0.5.\
But this is not possible since the max value of p should be 1.

## Problem 2 Bayesian network
a. DAG \
```{r}
dag <-
  model2network("[M][T][BC|M:T][HA|BC][S|BC][N|BC]")
graphviz.plot(dag)
```
\
b. What is the probability that a person will develop a blood clot as a result of both medication and trauma, and then have no medical implications? \
P(BC = T | M=T, T=T, N=T) \
= $\frac{P(BC=T, M=T, T=T, N=T)}{P(BC=T, M=T, T=T, N=T) + P(BC=F, M=T, T=T, N=T)}$\
= $\frac{P(M)P(T)P(BC|M,T)P(N|BC)}{P(M)P(T)P(BC|M,T)P(N|BC) + P(M)P(T)P(BC=F|M,T)P(N|BC=F)}$\
= $\frac{0.2 \cdot 0.05 \cdot 0.95 \cdot 0.25}{0.2 \cdot 0.05 \cdot 0.95 \cdot 0.25 + 0.2 \cdot 0.05 \cdot 0.05 \cdot 0.75)}$\
= $\frac{0.002375}{0.002375 + 0.000375}$ = 0.86363636...




## Problem 3
a. **What is the probability that you have the flu, given that it is winter, you have a headache, and you know that you are dehydrated?**\
P(F=T | H=T, D=T, S=winter)\
= $\frac{P(F,H,D,S)}{P(H,D,S)}$ \
= $\frac{P(S)P(F|S)P(D|S)P(H|F,D)}{P(S)P(F|S)P(D|S)P(H|F,D) + P(S)P(F=F|S)P(D|S)P(H|F=F,D)}$ \
= $\frac{0.5 \cdot0.4\cdot0.1\cdot0.9}{0.5 \cdot0.4\cdot0.1\cdot0.9 + 0.5 \cdot0.6\cdot0.1\cdot0.8}$ = 0.43\

b. **Does knowing you are dehydrated increase or decrease your probability of having the flu? Intuitively, does this make sense?** \
**Answer: Decrease.** It makes sense since knowing dehydrated may affect the probability of Headache.\
P(F=T | H=T, S=winter)\
= = $\frac{P(F,H,S)}{P(H,S)}$ \
= $\frac{P(S)P(F|S)P(D|S)P(H|F,D) + P(S)P(F|S)P(D=F|S)P(H|F,D=F)}{P(S)P(F|S)P(D|S)P(H|F,D) + P(S)P(F=F|S)P(D|S)P(H|F=F,D) + P(S)P(F|S)P(D=F|S)P(H|F,D=F) + P(S)P(F=F|S)P(D=F|S)P(H|F=F,D=F)}$ \
= $\frac{0.5 \cdot0.4\cdot0.1\cdot0.9 + 0.5 \cdot0.4\cdot0.9\cdot0.8}{0.5 \cdot0.4\cdot0.1\cdot0.9 + 0.5 \cdot0.6\cdot0.1\cdot0.8 + 0.5 \cdot0.4\cdot0.9\cdot0.8 + 0.5 \cdot0.6\cdot0.9\cdot0.3}$ = 0.61\
\
Compare P(F=T | H=T, S=winter) and P(F=T | H=T, D=T, S=winter), \
It implies that knowing dehydrated or not **decrease** the probability of having the flu. 

## Problem 4
```{r}
dag <-
  model2network("[A][C][F][B|A][D|A:C][E|B:F]")
graphviz.plot(dag)
```
\
a. Are A and F conditionally independent given B?\
Since node B won't imply any connection to node F, therefore it won't imply any connection between node A and F.\
So node A and F are **conditionally independent**.\
```{r}
dsep(bn = dag, x = "A", y = "F", 
     z = "B")
```


b. Are A and F conditionally independent given E? \
Since node E is one of the evidence node and both arrows lead into E,
it may provide connection between node A and F.\
So node A and F are **not conditionally independent**.\
```{r}
dsep(bn = dag, x = "A", y = "F", 
     z = "E")
```

c. Are A and F unconditionally independent?\
There will be no connection between A and F, so they would remain the same after 5 process taught in class.\
Therefore A and F are **independent.** (not conditionally independent)
```{r}
dsep(bn = dag, x = "A", y = "F")
```
